\documentclass{article}

\usepackage[final,nonatbib]{neurips_2019}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools, nccmath}
\usepackage[mathscr]{euscript}  % euler script font
\usepackage{cleveref}
\usepackage{enumitem}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

\usepackage{listings}
\usepackage{xcolor}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2
}

\lstset{style=mystyle}

\newtheorem{theorem}{Theorem}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\modexp}{mod\_exp}
\DeclareMathOperator{\decryptiontime}{decryption\_time}
\DeclareMathOperator{\extrareduction}{extra\ reduction}

\title{Timing Attacks against RSA}

\author{%
  Lorenzo Palloni\\
  University of Florence\\
  \texttt{lorenzo.palloni@stud.unifi.it} \\
}

\begin{document}

\maketitle

%----------------------------------------------------------------------------------------
\input{sections/introduction.tex}

%----------------------------------------------------------------------------------------
\input{sections/technical-background.tex}

%----------------------------------------------------------------------------------------
\input{sections/timing-attacks.tex}

%----------------------------------------------------------------------------------------
\input{sections/experiments.tex}

%----------------------------------------------------------------------------------------
\input{sections/defences.tex}

%----------------------------------------------------------------------------------------
\input{sections/conclusion.tex}

%----------------------------------------------------------------------------------------
% \section*{References}

\medskip
%\clearpage
{\small
\bibliographystyle{ieee}
\begin{thebibliography}{9}

\bibitem{bib:kocher}{Kocher, P.C., 1996, August. Timing attacks on implementations of Diffie-Hellman, RSA, DSS, and other systems. In Annual International Cryptology Conference (pp. 104-113). Springer, Berlin, Heidelberg.}
\bibitem{bib:openssl}{Brumley, D. and Boneh, D., 2005. Remote timing attacks are practical. Computer Networks, 48(5), pp.701-716.}
\bibitem{bib:boreale}{Boreale, M., 2003. Note per il corso di Sicurezza delle Reti.}
\bibitem{bib:montgomery}{Montgomery, P.L., 1985. Modular multiplication without trial division. Mathematics of computation, 44(170), pp.519-521.}
\bibitem{bib:sliding}{Menezes, A.J., Van Oorschot, P.C. and Vanstone, S.A., 2018. Handbook of applied cryptography. CRC press.}
\bibitem{bib:schindler}{Schindler, W., 2000, August. A timing attack against RSA with the chinese remainder theorem. In International Workshop on Cryptographic Hardware and Embedded Systems (pp. 109-124). Springer, Berlin, Heidelberg.}
\bibitem{bib:coppersmith}{Coppersmith, D., 1997. Small solutions to polynomial equations, and low exponent RSA vulnerabilities. Journal of cryptology, 10(4), pp.233-260.}

\end{thebibliography}
}
%----------------------------------------------------------------------------------------

\end{document}
